Sex and Power: sexual dimorphism in trait variability and its eco-evolutionary and statistical implications

### 2020-04-30 scripts/MouseSexDiffApril20_2020.Rmd
Re-organised the structure of the supplement to make it easier to follow the general structure, and so that Tables and Figures will match and be easier to find with the MS

# Set-up

## Loading packages & custom functions

## Packages

@@

## Functions 
Functions needed for preparing the data for meta analyses

@@

### 2) "Population statistics"
Create a function called: "calculate_population_stats"
This function groups animals from the same strain and institution together. This is done for each trait separately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

## Load & clean data

@@

### Loop: Meta-analyses on all traits

@@

### Merging datasets & removal of non-converged traits

@@

### Dealing with Correlated Parameters, preparation

@@

#### Collapsing and merging correlated parameters
Count of number of parameter names (correlated sub-traits) in each parameter group (par_group_size)

### Supplemental Table 2: "Numbers of correlated and uncorrelated traits"
This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

# Table detailing numbers of correlated and uncorrelated traits
kable(cbind(meta1 %>% count(parameter_group))) %>%
  kable_styling() %>%
  scroll_box(width = "60%", height = "200px")

@@

#### Meta-analyses on correlated (sub-)traits, using robumeta` 
Here we pepare the subset of the data (using nest()), and in this first step the model of the meta-analysis effect sizes are calculated

@@

### Performing meta-analyses 
Three for each of the 9 grouping terms: lnCVR, lnVR, lnRR

#### Preparation of data

@@

### Re-structuring the data for each grouping term
We here delete unused variables, and select the respective effect sizes. Please note - the referencing of the cells does NOT depend on previous ordering of the data. This would only be affected if the output structure from metafor::rma.uni changes.

@@

```
### Supplemental Table 4: "Meta-analysis on Functional Trait Groups"
SHINICHI
This table summarizes our overall meta-analyses on the functional trait group level ("GroupingTerm"). Results for lnCVR, lnVR and lnRR and their respective upper and lower 95% CI's, standard error and I2 values are provided.
```{r Table4}
kable(overall2, digits = 3) %>%
 kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

@@

# Visualisation
## Preparation for Figure 4 
#### Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

###  Overall results of second order meta-analysis (Figure 4, Panel B)
#### Re-structure data for plotting 
Data are re-structured, and grouping terms are being re-ordered

@@

### Figure 4

@@

## Preparation for Figure 5: Sex bias.
#### Preparing data for traits with CI not overlapping 0
To further investigate sex bias in this dataset, and in particular if the extent of sex bias differs between traits, we investigate the magnitude of male- and female bias in significantly different traits on (both for means and variability)

@@

###  Figure 5
Sex biases in trait values across trait groups (Step J in Figure 3). The panels in A show differences in the number of affected traits, in variance (lnCVR) and means (lnRR), where there is a significant difference between the sexes (i.e CI not overlapping zero). Panel B shows results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.

@@

# Supplemental Plots and Tables

### Supplemental Table S1: "Strains and Center Sample Sizes"
This table summarises the available numbers of male and female mice from each strain and originating institution.

```{r Table1_prep, include=FALSE}
# Initial checks
length(unique(data$parameter_name)) # 232 traits
length(unique(data$parameter_group)) # 161 parameter groups
length(unique(data$procedure_name)) # 26 procedure groups
length(unique(data$biological_sample_id)) # 27147 individial mice  
# To show a table with number of males and females per strain per production center:

ST1 <- (cbind(data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf))) 
```

```{r Table1}

@@

### Supplemental Table S3: "Meta-analysis results"

@@

### Supplemental Table S5: "Sex-biased effect sizes"
SHINICHI
This table provides some of the data plotted in Figures 5 and Figure S2. Here, we provide an overview over meta-analysis results performed on traits that were significantly biased to either sex. This table summarizes findings for both sexes and the respective functional trait groups.

@@

### Supplemental Table S6: "Heterogeneity within groups"
SHINICHI
This table summarizes our findings on heterogeneity due to institutions and mouse strains. These results are based on meta-analyses on sigma^2 and errors for mouse strains and centers (Institutions), following the identical workflow from above.

@@

### Figure S1 
#### Including lnVR
#### Count data, including lnVR (Fig S1 panel A)

The supplemental figures include lnVR (in addition to lnCVR and lnRR, as presented in Figures 4 and 5)
```{r FigS1_prep}
# *Prepare data for all traits

@@

### Heterogeneity
The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). 
As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

@@

### Figure S1

@@

### Preparations for Figure S2 A
This Figure extends Figure 4, as it includes results not only for lnCVR and lnRR but also lnCVR. In addition, we compare two different assessments of sex-bias, significance (CI not overlapping zero) and sex differences in male / female ratios > 10%
```{r}
# Plot FigS2 all significant results (CI not overlapping zero) for males, count data

@@

### Preparation for Figure S2 B (second-order meta-analysis)
#### Female Figure, significant and highly biased traits (Figure S2 B, left hand side panels)
Preparing data for traits with CI not overlapping 0: 
create column with 1= different from zero, 0= zero included in CI (for Figure S2 B, top left panel)
```{r FigS2_B}
#Restructure data for plotting

@@

# Metameta_Fig3_female.perc (Figure S2 bottom left panel)
```
#### Male Figure, significant and highly biased traits (Figure S2 B, right hand side panels)
Restructuring data with significant male bias (for Figure S2 B, top right panel)
```{r FigS2_B_right}
overall3.male.sigS <- gather(overall.male.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key = TRUE)

@@

### Figure S2

@@

## Acknowledgements
tbd

>>>>

# Set-up
## 1. Packages

@@

## 2. Functions 
Functions needed for preparing the data for meta analyses

@@

### 2) Population statistics
Create a function called: "calculate_population_stats"
This function groups animals from the same strain and institution together. This is done for each trait separately, and only for traits that have been measured in both sexes. Any group containing fewer than 5 individuals is excluded.

@@

## 3. Load & clean data

@@

### 1.1: Loop: Meta-analyses on all traits

@@

### 1.2: Merging datasets & removal of non-converged traits

@@

### 2.1: Dealing with Correlated Parameters, preparation

@@

### 2.2: Collapsing and merging correlated parameters, preparation for Table S2
Count the number of parameter names (correlated sub-traits) in each parameter group (par_group_size). This serves to identify and separate the traits that are correlated from the full dataset that can be processed as is. If the sample size (n) for a given "parameter group" equals 1, the trait is unique and uncorrelated. All instances, where there are 2 or more traits associated with the same parameter group (90 cases), are selected for a "mini-meta analysis", which removes the issue of correlation.

@@

@@

### 2.3: Meta-analyses on correlated (sub-)traits, using `robumeta` 
Here we pepare the subset of the data (using nest()), and in this first step the model of the meta-analysis effect sizes are calculated

@@

### 3.1: Performing meta-analyses 
Three for each of the 9 grouping terms: lnCVR, lnVR, lnRR
#### 3.1.1: Preparing the data

@@

#### 3.1.2: Re-structuring the data for each grouping term
We here delete unused variables, and select the respective effect sizes. Please note - the referencing of the cells does NOT depend on previous ordering of the data. This would only be affected if the output structure from metafor::rma.uni changes.

@@

@@

# Visualising meta-analysis results
## 1. Preparation for Figure 4 
#### 1.1: Count data, based on First-order meta-analysis results
This includes all separate eligible traits.
Re-ordering of grouping terms. 
Preparation of subplot 4A

@@

#### 1.2: Overall results of second order meta-analysis (Figure 4, Panel B)
#### 1.3: Re-structure data for plotting 
Data are re-structured, and grouping terms are being re-ordered

@@

## 2. Figure 4

@@

## 3. Preparation for Figure 5
#### 3.1: Preparing data for traits with CI not overlapping 0
To further investigate sex bias in this dataset, and in particular if the extent of sex bias differs between traits, we investigate the magnitude of male- and female bias in significantly different traits on (both for means and variability)

@@

## 4. Figure 5
Sex biases in trait values across trait groups (Step J in Figure 3). The panels in A show differences in the number of affected traits, in variance (lnCVR) and means (lnRR), where there is a significant difference between the sexes (i.e CI not overlapping zero). Panel B shows results for the sex bias in those traits that differ between the sexes (second-order meta-analysis). Triangles represent sex bias in means (response ratio) and black circles differences in the coefficient of variation ratio (mean-adjusted variability). The orange-red bars represent trait groups with a female bias, blue-green bars male-biased traits.

@@

# Supplemental Tables and Figures

## Supplemental Tables
#### 1. Supplemental Table S1: "Strains and Center Sample Sizes"
  This table summarises the available numbers of male and female mice from each strain and originating institution.

```{r Table1_prep}
  # Initial checks
  length(unique(data$parameter_name)) # 232 traits
  length(unique(data$parameter_group)) # 161 parameter groups
  length(unique(data$procedure_name)) # 26 procedure groups
  length(unique(data$biological_sample_id)) # 27147 individial mice  
  # To show a table with number of males and females per strain per production center:

  ST1 <- (cbind(data %>% group_by(production_center, strain_name) %>% count(biological_sample_id, sex) %>% count(sex) %>% print(n = Inf))) 
```
```{r Table1}

@@

#### 2. Supplemental Table S2: "Numbers of correlated and uncorrelated traits"

```{r, TableS2}  
# Table detailing numbers of correlated and uncorrelated traits
kable(cbind(meta1 %>% count(parameter_group))) %>%
  kable_styling() %>%
  scroll_box(width = "60%", height = "200px")
```

#### 3. Supplemental Table S3: "Meta-analysis results"

@@

#### 4. Supplemental Table S4: "Meta-analysis on Functional Trait Groups"
  SHINICHI
  This table summarizes our overall meta-analyses on the functional trait group level ("GroupingTerm"). Results for lnCVR, lnVR and lnRR and their respective upper and lower 95% CI's, standard error and I2 values are provided.
```{r Table4}
  kable(overall2, digits = 3) %>%
   kable_styling() %>%
  scroll_box(width = "100%", height = "200px")
```

#### 5. Supplemental Table S5: "Sex-biased effect sizes"
  SHINICHI
  This table provides some of the data plotted in Figures 5 and Figure S2. Here, we provide an overview over meta-analysis results performed on traits that were significantly biased to either sex. This table summarizes findings for both sexes and the respective functional trait groups.

@@

#### 6. Supplemental Table S6: "Heterogeneity within groups"
  SHINICHI
  This table summarizes our findings on heterogeneity due to institutions and mouse strains. These results are based on meta-analyses on sigma^2 and errors for mouse strains and centers (Institutions), following the identical workflow from above.

@@

## Supplemental Figures
### 1. Preparation for Figure S1
#### 1.1: Preparing the data
The supplemental figures include lnVR (in addition to lnCVR and lnRR, as presented in Figures 4 and 5)
```{r FigS1_prep}
# *Prepare data for all traits

@@

#### 1.2: Preparing heterogeneity
The analysis for heterogeneity follows the workflow of the above steps for the different meta-analyses. However, in the initial meta-analysis we extract sigma^2 and errors for mouse strains and centers (Institutions). 
As above a Loop is run and parameters extracted from metafor (sigma2's, s.nlevels)

@@

### 2. Figure S1

@@

### 3. Preparation for Figure S2
#### 3.1: Preparation for Figure S2 A
This Figure extends Figure 4, as it includes results not only for lnCVR and lnRR but also lnCVR. In addition, we compare two different assessments of sex-bias, significance (CI not overlapping zero) and sex differences in male / female ratios > 10%
```{r}
# Plot FigS2 all significant results (CI not overlapping zero) for males, count data

@@

#### 3.2: Preparation for Figure S2 B (second-order meta-analysis)
#### 3.3: Female Figure, significant and highly biased traits (Figure S2 B, left hand side panels)
Preparing data for traits with CI not overlapping 0: 
create column with 1= different from zero, 0= zero included in CI (for Figure S2 B, top left panel)
```{r FigS2_B}
#Restructure data for plotting

@@

# Metameta_Fig3_female.perc (Figure S2 bottom left panel)
```
#### 3.4: Male Figure, significant and highly biased traits (Figure S2 B, right hand side panels)
Restructuring data with significant male bias (for Figure S2 B, top right panel)
```{r FigS2_B_right}
overall3.male.sigS <- gather(overall.male.plot3, parameter, value, c(lnCVR, lnVR, lnRR), factor_key = TRUE)

@@

### 4. Figure S2

@@
